---
title: "Final Project"
author: "Megha Khanna"
date: "04/15/2024"
---

#For my Final project, I have utilized various codes to analyze a dataset of newspaper publications concerning Jack the Ripper and his notorious murders.

My research questions are as follows:

1.  How does the frequency and context of terms associated with Jack the Ripper differ from those related to his victims in historical newspaper articles, and what does this tell us about societal attitudes towards crime and victimhood in late Victorian England?

2.  How did the topics covered in newspaper reports change before and after 1888, the year of Jack the Ripper's crimes? What can this reveal about the influence of these events on public discourse and media focus?

##Introduction

The infamous Jack the Ripper murders of 1888, targeting several sex workers in London’s Whitechapel district, became a stark illustration of the interplay between media sensationalism, public fear, and the stigmatization of sex workers. Jack the Ripper’s name continues to resonate throughout criminal history, not just for the horrific nature of the crimes but also for the media frenzy that erupted in late Victorian England.

This analysis aims to explore the differential portrayal of Jack the Ripper compared to his victims through the lens of historical newspaper articles. By analyzing the frequency and context of words used, we can gain insights into the narrative and societal perceptions constructed by the media of that era. The other half of the analysis looks at the pre and post year 1888 and how the narratives related to crimes have significantly changes with the sensationalism of ripper crimes.

```{r}
#1 Loading libraries
library(tm)
library(topicmodels)
library(tidytext)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)

```

```{r}
#2 Reading data
casebook_final_2 <- read_csv("~/Downloads/casebook/casebook final 2.csv", show_col_types = FALSE)

```

```{r}
#3 Defining stopwords that were coming often in my dataset and then combining them with custom stopwords.
custom_stopwords <- c("aberdeen", "weekly", "times", "journal", "article", "said")
all_stopwords <- c(stopwords, custom_stopwords)

```

```{r}
#4 Now I am preprocessing the text and tokenizing
tidy_casebook <- casebook_final_2 %>%
  unnest_tokens(word, News.Reports) %>%                 
  filter(!word %in% all_stopwords) %>%                  
  mutate(word = gsub("[\\d-]+", "", word)) %>%         
  anti_join(data.frame(word = stopwords::stopwords(language = "en", source = "snowball")), by = "word")  

```

```{r}
#5 Defining keywords and filter texts for ripper and sex workers
keywords_jack <- c("jack the ripper", "jack", "murderer", "villain", "madman", "mystery", 
                   "killer", "ripper", "assassin", "butcher", "bogey man", "Whitechapel Murderer", "leather Apron")
keywords_victims <- c("victim", "sex worker", "prostitute", "deceased", "body", "woman", 
                      "women", "lady", "ladies", "alone", "public woman", "public", 
                      "public women", "dead woman", "corpse", "girl", "she", "her", 
                      "ill repute", "streetwalker", "harlot", "fallen")


```

```{r}
#6 Counting occurrences of each
jack_texts <- tidy_casebook %>%
  filter(str_detect(word, paste(keywords_jack, collapse = "|")))
victim_texts <- tidy_casebook %>%
  filter(str_detect(word, paste(keywords_victims, collapse = "|")))

jack_count <- nrow(jack_texts)
victim_count <- nrow(victim_texts)

```

```{r}
#7 Word frequency analysis for each and combined
jack_word_freq <- jack_texts %>%
  count(word, sort = TRUE) %>%
  top_n(10, n) %>%
  mutate(frequency = n / sum(n))

victim_word_freq <- victim_texts %>%
  count(word, sort = TRUE) %>%
  top_n(10, n) %>%
  mutate(frequency = n / sum(n))

combined_top10_freq <- full_join(jack_word_freq, victim_word_freq, by = "word", suffix = c("_jack", "_victim"))

jack_word_freq$dataset <- "Jack" 
victim_word_freq$dataset <- "Victim"

```

```{r}
#8 Combining datasets and plotting
combined_top10_freq <- bind_rows(jack_word_freq, victim_word_freq)

ggplot(combined_top10_freq, aes(x = reorder(word, frequency), y = frequency, fill = dataset)) +
  geom_col() +
  facet_wrap(~ dataset, scales = "free_y") +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Words", y = "Frequency", title = "Top 10 Frequencies of Words Related to Jack the Ripper and Victims")

```

According to the results, I found that the disparity in the frequency of terms between Jack the Ripper and his victims could highlight a potential imbalance in media coverage. The media may have focused more on the sensational aspects of the murderer's identity rather than the victims' sories. The context in which these terms appear (aggressive and mysterious for the Ripper, generic and impersonal for the victims) suggests a narrative that prioritizes the sensationalism of the criminal over the humanization of the victims. This can be a reflection of societal attitudes where the story of the criminal is more engrossing than the plight of the marginalized. This confirms Warkentin's observation where she notes the glorification of the murderer and the trivialization of his female victims. She points out that the original narrative painted the Ripper as a nearly heroic figure, while his victims, mostly sex workers, were depicted in derogatory terms. This narrative served to enforce societal norms and control women's behavior​​.

The language used indicates a society fascinated by the macabre details of the crimes and the enigmatic figure of the Ripper, while the victims are portrayed in a way that may perpetuate stigmas, particularly against women in vulnerable positions. The data suggests a societal tendency to sensationalize criminals, which could detract from a compassionate and just representation of victims, particularly of sex workers in this context.

```{r}
#1 Loading libraries
library(tidytext)
library(topicmodels)
library(lubridate)
library(dplyr)
library(tm)
library(stringr)

```

```{r}

```
